\Question{Training and Validation Error}
Suppose you have split your data into training and validation sets and have
trained your model on the training set. You are now
examining its training and validation errors. What could be going on in each of these cases below and what
might we try to decrease the high errors?

\begin{Parts}
    \Part High training error, high validation error

    \begin{solution}
        The model is underfitting and is unable to capture the patterns in
        either the training or validation set. We might try switching to a more
        complex model with greater expressiveness to solve this.
    \end{solution}

    \Part High training error, low validation error

    \begin{solution}
        The way our data was split was unlucky for us, with a training set with
        particularly "hard" cases to learn and a validation set with
        particularly "easy" cases to predict. We might try gathering more data
        as in part (a), or use k-fold cross validation or another
        training/validation splitting method that gives all the data a chance
        to be in both the validation and training sets.
    \end{solution}

    \Part Low training error, high validation error

    \begin{solution}
        The model is overfitting to the training data. We might try adding
        regularization methods, using a simpler model, stopping the training
        early, selecting only certain features to use in the model to reduce
        the validation error.
    \end{solution}

    \Part Low training error, low validation error

    \begin{solution}
        The model is performing well and we are doing good! We might try coming
        up with a new breakthrough state of the art model that captures
        the patterns in the data even better.
    \end{solution}
\end{Parts}
